---
title: "20-soc-analysis"
output: html_notebook
---
#soc field grouped by year
```{r}
h1b_df %>% 
  filter(year == 2018)%>% 
  group_by(soc_field) %>% 
  summarise(n = n())  %>% 
  arrange(-n)

top_6_soc <- h1b_df %>% 
  group_by(soc_field, year) %>% 
  summarise(n = n()) %>% 
  filter(soc_field %in% c(15,13,17,11,19,29)) %>% 
  ungroup()

```
soc fields 15, 13, 17, 11 and 19 have had the most aplications
    15-0000 Computer and Mathematical Occupations
    13-0000 Business and Financial Operations Occupations
    17-0000 Architecture and Engineering Occupations
    11-0000 Management Occupations
    19-0000 Life, Physical, and Social Science Occupations
    29-0000 Healthcare Practitioners and Technical Occupations
    
    
```{r}
top_6_soc$soc_name[top_6_soc$soc_field == 15] <- 'Computer and Mathematical Occupations'
top_6_soc$soc_name[top_6_soc$soc_field == 13] <- 'Business and Financial Operations Occupations'
top_6_soc$soc_name[top_6_soc$soc_field == 17] <- 'Architecture and Engineering Occupations'
top_6_soc$soc_name[top_6_soc$soc_field == 11] <- 'Management Occupations'
top_6_soc$soc_name[top_6_soc$soc_field == 19] <- 'Life, Physical, and Social Science Occupations'
top_6_soc$soc_name[top_6_soc$soc_field == 29] <- 'Healthcare Practitioners and Technical Occupations'

```

    

#Analysis of all other except for 15
```{r}
top_6_soc %>% 
  filter(!soc_field== 15) %>% 
  ggplot(aes(x = year, y = n))+
  geom_line(aes(color = soc_name), size = 1.05)+
  labs( y = "Number of Applications", x = "Year")+
  theme_bw()
```


#Analysis of 15 by itself
```{r}
top_6_soc %>% 
  filter(soc_field == 15) %>% 
  ggplot(aes(x = year, y = n))+
  geom_line(color = 'red', size = 1.15)+
  theme_bw()
```

#Analysis of top job titles in soc == 15
```{r}
soc_tech_14 <- h1b_df %>% 
  filter(year == 2014) %>% 
  filter(soc_field == 15) %>% 
  group_by(employer_name) %>% 
  summarise(n = n()) %>% 
  arrange(-n) %>% 
  ungroup

soc_tech_18 <- h1b_df %>% 
  filter(year == 2018) %>% 
  filter(soc_field == 15) %>% 
  group_by(employer_name) %>% 
  summarise(n = n()) %>% 
  arrange(-n) %>% 
  ungroup

soc_tech_16 <- h1b_df %>% 
  filter(year == 2018) %>% 
  filter(soc_field == 15) %>% 
  group_by(employer_name) %>% 
  summarise(n = n()) %>% 
  arrange(-n) %>% 
  ungroup

soc_tech_year <- h1b_df %>% 
  filter(soc_field == 15) %>% 
  group_by(employer_name, year) %>% 
  summarise(n = n()) %>% 
  arrange(-n) %>% 
  ungroup

soc_tech
```

```{r}
soc_10_14 <- soc_tech_14 %>% 
  top_n(15)

soc_10_18 <- soc_tech_18 %>% 
  top_n(15)

soc_top <- as.vector(intersect(soc_10_14$employer_name,soc_10_18$employer_name))
soc_top

```

#Analysis of how the #of applications for top companies has changed

```{r}
soc_tech_year %>% 
  filter(employer_name %in% soc_top) %>% 
  ggplot(aes(x = year, y = n))+
  geom_line(aes(color = employer_name), size = 1.05)+
  labs( y = "Number of Applications", x = "Year")+
  theme_bw()
  
```

#Top Tech Companies
```{r}
#Apple
soc_tech_year %>% 
  filter(str_detect(employer_name, 'apple inc')) %>% 
  mutate(employer_name = 'Apple')

#Microsoft
soc_tech_year %>% 
  filter(str_detect(employer_name, 'microsoft')) %>% 
  mutate(employer_name = 'Microsoft')

#Samsung
soc_tech_year %>% 
  filter(str_detect(employer_name, 'samsung')) %>% 
  mutate(employer_name = 'Samsung')

#Intel
soc_tech_year %>% 
  filter(str_detect(employer_name, 'intel corp|intel mob|intel ame|intel mas|intel fed')) %>% 
  mutate(employer_name = 'Intel')

#Google
soc_tech_year %>% 
  filter(str_detect(employer_name, 'google'))%>% 
  mutate(employer_name = 'Google')
  
#Facebook
soc_tech_year %>% 
  filter(str_detect(employer_name, 'facebook')) %>% 
  mutate(employer_name = 'Facebook')

#IBM
soc_tech_year %>% 
  filter(str_detect(employer_name, 'ibm')) %>% 
  mutate(employer_name = 'IBM')

soc_tech_top <- soc_tech_year %>% 
  mutate(employer_name = ifelse(str_detect(employer_name,'apple inc'), 'Apple',
                          ifelse(str_detect(employer_name, 'microsoft'), 'Microsoft',
                          ifelse(str_detect(employer_name, 'samsung'), 'Samsung',
                          ifelse(str_detect(employer_name,'intel corp|intel mob|intel ame|intel mas|intel fed'), 'Intel', 
                          ifelse(str_detect(employer_name,'facebook'),'Facebook',
                          ifelse(str_detect(employer_name,'google'),'Google',
                          ifelse(str_detect(employer_name, 'ibm'),'IBM',"no")))))))) 


         
         
soc_tech_top %>% 
  group_by(employer_name, year) %>% 
  summarise(n = n()) %>% 
  filter(!employer_name == "no") %>% 
  ungroup() %>% 
  ggplot(aes(x = year, y = n))+
  geom_line(aes(color = employer_name), size = 1.05)+
  labs( y = "Number of Applications", x = "Year")+
  theme_bw()
```

